---
title: "Unit1-IntrotoAnalytics"
author: "Raka"
date: "9/21/2020"
output: html_document
---

## Getting started with R: Vector and dataframe

I would like to document all of my learning journey for this specific course on EDX: MITx Analytics Edge, offered to students who are interesting in using R as the data science engine. For those who are interested, go to: https://www.edx.org/course/the-analytics-edge for details.

```{r vector&dataframe}
c(2,5,1,6,3)
country <- c("Brazil", "Indonesia", "China", "USA")
country
lifeExpectancy <- c(78,75,72,81)

# Getting specific element. ex: Indonesia, number 2
country[2]

```

## Creating sequence of numbers 

We can also create a sequence of numbers using seq() function. See below.

```{r sequence}
seq(0,100,2)
```

## Data frame in R

```{r dataframe}
countryData <- data.frame(country, lifeExpectancy)
countryData
```
## Adding data into the dataframe

Bear in mind, I just randomly created these numbers without checking their actual values.

```{r addDF}
countryData$population <- c(10000,22222,64545454,11101)
countryData
```

## Combining/ appending dataframe

For instance, we have 2 extra countries -- to be addded as rows.

```{r newrows}
country <- c("Switzerland", "France")
lifeExpectancy <- c(88,81)
population <- c(9000000,1191919)

newCountryData <- data.frame(country, lifeExpectancy, population)

# Adding this df to old df
AllCountrydata <- rbind(countryData, newCountryData)
AllCountrydata
```
## Reading data from a CSV file

Let's now read a CSV file in our folder. Check your directory to see the generated file :).

```{r readCSV}
WHO <- read.csv("WHO.csv")
WHO
str(WHO)

# We can even call a summary function
summary(WHO)
```

## Taking only a small subset of DF

What about taking smaller fractions of our data? Yes we can, using subset function.
Afterward, we output this into "WHO_Europe.csv" file.

```{r subset}
WHO_Europe <- subset(WHO, Region == "Europe")
str(WHO_Europe)

# Storing this subset into a separate csv
write.csv(WHO_Europe, "WHO_Eur.csv")

# This line is used to check our variables
ls()

# We can also remove a specific variable by using rm()
rm(countyrdata) #terribly misspelled

```

# Doing basic data analysis

We can practice some data analysis in this exercise such as mean, etc.

```{r analyses}
WHO$Under15
mean(WHO$Under15)
# and std dev

sd(WHO$Under15)

# or specific variable
summary(WHO$Under15)
```

## Can we index to a specific value? YES!

For instance, can we identify, which country has the the lowest or highest value in Under15?
Japan is the answer.

```{r whichmin}
which.min(WHO$Under15)
which.max(WHO$Under15)
# Now we use this as indexing

WHO$Country[which.min(WHO$Under15)]
WHO$Country[which.max(WHO$Under15)]
```

## Plotting world in R

So, let's start with the scatter plot: the mother of all plots.

```{r scatter}
plot(x = WHO$GNI, y = WHO$FertilityRate)
```
## Managing outliers in dataframe, after seeing the plot
```{r}

```

Can we spot the outliers from figure above? yep, of course!

```{r outliers}
outliers <- subset(WHO, GNI > 10000 & FertilityRate > 2.5)
nrow(outliers) 

# Then we output them
outliers[c("Country", "GNI", "FertilityRate")]

```
## Answering some questions on the website

These are just some lines to answer questions after the video. Feel free to skip.

```{r answers}
colnames(WHO)
mean(WHO$Over60)
WHO$Country[which.min(WHO$Over60)]
WHO$Country[which.max(WHO$LiteracyRate)]
```

# Fun parts: Playing with statistical distributions on R.

Here comes the interesting journey in R, how easy it is to build plots (somewhat complex in EXCEL) with only 1 - 2 lines of code.
Embrace!

```{r distr}
hist(WHO$CellularSubscribers)
boxplot(WHO$LifeExpectancy ~ WHO$Region)

# we fix some labelling and title.
boxplot(WHO$LifeExpectancy ~ WHO$Region, 
        xlab = "", ylab = "Life Expectancy",
        main = "Life Expectancy of Countries by Region")

```
# Creating a summary output in a table

We can also create a table from our dataframe as the summary output.

```{r summary}
table(WHO$Region)

```

# Can we use it to find aggregated results? Ex: the percentage of pop over 60 in these regions. HOW?

Yes, we use the function called tapply :) Behold.

```{r tappl}
tapply(WHO$Over60, WHO$Region, mean)

# Lets then find other results
tapply(WHO$LiteracyRate, WHO$Region, min, na.rm = TRUE)
```

## Answering questions from the course, part 2

If you like, you can skip again this part.

```{r answers2}
tapply(WHO$ChildMortality, WHO$Region, min, na.rm = TRUE)
```

## The start of Recitation: USDA Nutritional Education, Understanding food

Allright folks, now we enter the recitation part of this week. Are you ready?
Yes, I am so pumped too. Let's get started!
Of course, the first thing to do is to load the data

```{r loadUSDA}
USDA <- read.csv("USDA.csv")

#and somwewhat mandatory steps in any EDA, Str and Summary
str(USDA)
summary(USDA)
```

## There is a weird number!

The amount of sodium (max value) that summary function gives us is way too high.
Let's find out what that is.

```{r sodium}
names(USDA)
USDA$Description[which.max(USDA$Sodium)]

```

# What about high containing sodium?

We can create a subset of data, using subset function indeed :)

```{r sodsub}
HiSodium <- subset(USDA, Sodium > 10000)
nrow(HiSodium)
HiSodium$Description
```

## Finding the sodium value of "CAVIAR"

Yes, we can also filter using 'string' as input command.

```{r caviar}
match("CAVIAR", USDA$Description)
USDA$Sodium[match("CAVIAR", USDA$Description)]

# descriptive stas
mean(USDA$Sodium, na.rm = T)
sd(USDA$Sodium, na.rm = T)
```

## Entering the visualization part of week 1 USDA DATA

Here we go, my favorite part totally!

```{r viz}
plot(USDA$Protein, USDA$TotalFat)
plot(USDA$Protein, USDA$TotalFat, xlab = "Protein", ylab = "Fat",
     main = "Protein vs Fat", col = "red")

# Histogram of VItC Levels
hist(USDA$VitaminC, xlab = "Vit C (mg)", main = "Histogram of Vit C level")
hist(USDA$VitaminC, xlab = "Vit C (mg)", main = "Histogram of Vit C level", xlim = c(0,100), breaks = 2000)
```

## Plotting in boxplot

We look at the sugar level now :)

```{r bxplt}
boxplot(USDA$Sugar, main = "Boxplot of Sugar Levels", ylab = "Sugar (g)")
```
## How about creating variables using logical data as results? Sure thing bro

We can assign our logical results as numeric. See below.

```{r variables}
HighSodium <- USDA$Sodium > mean(USDA$Sodium, na.rm = TRUE)
HighSodium

# Assign them as numeric results: 1 or 0

HighSodium <- as.numeric(USDA$Sodium > mean(USDA$Sodium, na.rm = TRUE))
HighSodium

# Then insert this into a new column
USDA$HiSodium <- as.numeric(USDA$Sodium > mean(USDA$Sodium, na.rm = TRUE))
str(USDA)

# and do the same for other variables
USDA$HiProtein <- as.numeric(USDA$Protein > mean(USDA$Protein, na.rm = TRUE))
USDA$HiFat <- as.numeric(USDA$TotalFat > mean(USDA$TotalFat, na.rm = TRUE))
USDA$HiCarbs <- as.numeric(USDA$Carbohydrate > mean(USDA$Carbohydrate, na.rm = TRUE))
str(USDA)
```

## Creating a summary table from our previous results

We can now summarize our analysis and present them in a chart.

```{r table2}
table(USDA$HiSodium, USDA$HiFat)
```

## Looking for specifics: Food that contains iron, can we calculate their average iron contents based on two groups of Hi Protein?

Yes, why not!
We use tapply for that.

```{r tapp}
tapply(USDA$Iron, USDA$HiProtein, mean, na.rm = TRUE)

# With the same approach, to find max vit C between hi carb/ lo carb food
tapply(USDA$VitaminC , USDA$HiCarbs, max, na.rm = TRUE)
# we can even use summary!
tapply(USDA$VitaminC , USDA$HiCarbs, summary, na.rm = TRUE)
```

