---
title: "MITx_MoneyballWeek2"
author: "Raka"
date: "10/4/2020"
output: html_document
---

## Moneyball, baseball game

We first load the dataset and build the regression model here.

```{r loadbase}
baseball <- read.csv("baseball.csv")
str(baseball)

moneyball <- subset(baseball, Year < 2002)
str(moneyball)

moneyball$RD <- moneyball$RS - moneyball$RA
str(moneyball)

plot(moneyball$RD, moneyball$W)
WinsReg <- lm(W ~ RD, data = moneyball)
summary(WinsReg)

# W = 80.88 + 0.105*(RD)
# 80.88 + 0.1058(RD) >= 95, hence RD >= 135 =)
```

# Quiz 1 Question

If a baseball team scores 713 runs and allows 614 runs, how many games do we expect the team to win?

Using the linear regression model constructed during the lecture, enter the number of games we expect the team to win:

```{r quiz1}

RD_1 <- 713 - 614
# Thus,

W_1 <- 80.88 + 0.1058*RD_1
W_1

```

# We build RunsReg based on OBP, SLG and BA

```{r runs}
str(moneyball)

RunsReg <- lm(RS ~ OBP + SLG + BA, data = moneyball)
RunsReg2 <- lm(RS ~ OBP + SLG, data = moneyball)

# OOBP on base percentage, SLG slugging percentage
summary(RunsReg2)
```

# If a baseball team's OBP is 0.311 and SLG is 0.405, how many runs do we expect the team to score?

Using the linear regression model constructed during the lecture (the one that uses OBP and SLG as independent variables), enter the number of runs we expect the team to score: (1)

If a baseball team's opponents OBP (OOBP) is 0.297 and oppenents SLG (OSLG) is 0.370, how many runs do we expect the team to allow?

```{r predictruns}

# RegRuns = -804.6 + 2737.7*OBP + 1584.9*SLG
OBP_1 <- 0.311
SLG_1 <- 0.405
Runs <- -804.6 + 2737.7*OBP_1 + 1584.9*SLG_1

Runs

# RunsAllowed in Video

RA_1 <- -837.4 + 2913.6*0.297 + 1514.3*0.37
RA_1
```

# Quick Question

In 2012 and 2013, there were 10 teams in the MLB playoffs: the six teams that had the most wins in each baseball division, and four "wild card" teams. The playoffs start between the four wild card teams - the two teams that win proceed in the playoffs (8 teams remaining). Then, these teams are paired off and play a series of games. The four teams that win are then paired and play to determine who will play in the World Series. 

We can assign rankings to the teams as follows:

    Rank 1: the team that won the World Series
    Rank 2: the team that lost the World Series
    Rank 3: the two teams that lost to the teams in the World Series
    Rank 4: the four teams that made it past the wild card round, but lost to the above four teams
    Rank 5: the two teams that lost the wild card round

In your R console, create a corresponding rank vector by typing

```{r qques}

teamRank = c(1,2,3,3,4,4,4,4,5,5)

wins2012 <- c(94,88,95,88,93,94,98,97,93,94)
wins2013 <- c(97,97,92,93,92,96,94,96,92,90)

cor(teamRank, wins2012)
cor(teamRank, wins2013)

```

